---
title: "Untitled"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
bibliography: ../data/r4biochem.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

# Background
```{r importdata}
rawcount <- read_csv("../data/rawcounts.csv")
#head(rawcount)
```

we sent 10 samples to XYZ lab for expression profiles from NUMBER2 genes. the file is returned with the following column names ```r colnames(rawcount)```. We used the methods from [@breitling2004] in our analysis and maybe use the bioconductor package [@gentleman2004]


# Summary

```{r tidy}

## Tidying
gene_dat <- rawcount %>% 
  # separate the sgRNA column into multiple new columns based on the "_" character as a delimiter 
    separate(col = sgRNA, into = c("something","TSS", "strand", "probe_gene"), sep = "_") %>%
  
  # separate the column that has the probe and gene info into two columns based on the "-" character
    separate(col = probe_gene, into = c("probe", "gene"), sep = "-") %>% 
  
  ## mutate
  # Calculate fold change/ probe by using T20/T0
  mutate(fold_change = T20 / T0)

```

Table of Gene Summary Data
```{r genetable}
## gene summary
gene_summary <- gene_dat %>%
    group_by(gene) %>%
    summarise(mean = mean(fold_change, na.rm = T), 
              sd = sd(fold_change, na.rm = T),
              n_probes = n())

knitr::kable(gene_summary, 
             digits = 3, # number of digits
             align = "lccc", # column alignment
             caption = "Summary fold change"
             )
```

Plot Fold Change by Gene

```{r geneplot, fig.height = 10, fig.align = "center", echo = FALSE }
p1 <- ggplot(gene_dat, aes(x = gene, y = fold_change)) +
  #geom_boxplot()
  geom_violin(alpha = 0) +
  geom_jitter(alpha = 0.6, width = 0.1, color = "blue") +
  labs(title = "Distribution of Fold Change\nby Gene",
       x = "Fold Change",
       y = "Gene Name") +
  theme(axis.text.x  = element_text(angle=45, vjust=0.6))
  

p1
```


# Discussion

```{r bestgene}
best <- gene_summary %>% 
  filter(n_probes == max(n_probes)) # filter to the gene with most probes
```

For no better reason than it's a tidy looking name and has the most probes (n = `r best$n_probes`), we are interested in the gene ***`r best$gene`*** 

# References
